---
title: "Basics"
author: "Valerio Villani"
date: "30/03/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, error=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)

theme_set(theme_bw())

```

# The Ames housing data

## Goal

Predict the sale price of a house based on its other characteristics.

```{r echo=FALSE, include=FALSE}

data(ames, package = "modeldata")

# ames <- janitor::clean_names(ames)

glimpse(ames)

```

## EDA

```{r echo=FALSE}

ames %>% 
  ggplot(aes(x = Sale_Price)) +
  geom_histogram(bins = 50) +
  # visualise log-transformed data
  scale_x_log10()


```
```{r include=FALSE}

# apply log-transform to Sale_Price
ames <- ames %>% mutate(Sale_Price = log10(Sale_Price))

```


# Spending our data

The *data spending* stage is where we decide the proportion of our **train/test split**.

## Common methods for splitting data

```{r}

set.seed(123)

# calculate quantiles
qtls <- quantile(ames$Sale_Price, 
                 probs = c(0.25, 0.50, 0.75))

# visualise quantiles for stratified sampling
ames %>% 
  ggplot(aes(x = Sale_Price)) +
  geom_density() +
  geom_vline(xintercept = qtls, linetype = "dashed" ) +
  labs(x = "Sale price (log-10 USD)")



# split data into 80% train and 20% test
ames_split <- initial_split(ames, prop = 0.80,
                            # sample according to quartiles
                            strata = Sale_Price)

# get the train data
ames_train <- training(ames_split)

# get the test data
ames_test <- testing(ames_split)

```

# Feature engineering

Feature engineering encompasses activities that reformat predictor values to make them easier for a model to use effectively. One of the most common feature engineering tasks is transforming nominal or qualitative data (factors or characters) so that they can be encoded or represented numerically.


```{r}

lm(Sale_Price ~ Neighborhood + log10(Gr_Liv_Area) + Year_Built + Bldg_Type,
   data = ames)

```
Remember that when invoking the `recipe()` function, the steps are not estimated or executed in any way. The second phase for using a recipe is to estimate any quantities required by the steps using the `prep()` function.

```{r recipe}

# create a lm() recipe
ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>% 
  # log-transform "general living area" variable
  step_log(Gr_Liv_Area, base = 10) %>% 
  # dummy-code all factors
  step_dummy(all_nominal())

ames_rec


```

```{r prep}

simple_ames <- prep(ames_rec, training = ames_train)

simple_ames

```
```{r bake}

ames_test_bake <- bake(simple_ames, new_data = ames_test)

names(ames_test_bake) %>% 
  head()

```

```{r}

# we have got too many neighbourhoods
ggplot(ames_train, aes(y = Neighborhood)) + 
  geom_bar() + 
  labs(y = NULL)

```
```{r recipe2}

ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  # lump the bottm 1% of neighbourhoods into the "other" category
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal())

ames_rec

```

After exploring the Ames training set, we might find that the regression slopes for the general living area differ for different building types.

```{r interaction-terms}

ggplot(ames_train, aes(x = Gr_Liv_Area, y = 10^Sale_Price)) + 
  geom_point(alpha = .2) + 
  facet_wrap(~Bldg_Type) + 
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, col = "red") + 
  scale_x_log10() + 
  scale_y_log10() + 
  labs(x = "General Living Area", y = "Sale Price (USD)")

```


```{r recipe-interaction}

ames_rec <- recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
                   data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal()) %>% 
  # Gr_Liv_Area is on the log scale from a previous step
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_type_") )

ames_rec

```

## Spline functions

```{r}

library(patchwork)
library(splines)

plot_smoother <- function(deg_free) {
  ggplot(ames_train, aes(x = Latitude, y = Sale_Price)) + 
    geom_point(alpha = .2) + 
    scale_y_log10() +
    geom_smooth(
      method = lm,
      formula = y ~ ns(x, df = deg_free),
      col = "red",
      se = FALSE
    ) +
    ggtitle(paste(deg_free, "Spline Terms"))
}

( plot_smoother(2) + plot_smoother(5) ) / ( plot_smoother(20) + plot_smoother(100) )

```
The `ns()` function in the splines package generates feature columns using functions called natural splines.


```{r recipe-nonlinear}

recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  # apply natural splines to Latitude
  step_ns(Latitude, deg_free = 20)

```

